Quantitative Auditing of Fairness Measures with Differentially Private Synthetic Data

### 2024-12-01 paper/paper.tex
reorg

After synthesizing datasets, we used the fairness checker from \cite{yuan2024ensuring} to compute their fairness measures. To test the viability of our method, we compare the metrics computed from the synthetic dataset against those of the original dataset. We used various datasets with fairness concerns mentioned in \cite{pessach2022review}.

@@

TODO: section experiment vs Methodology. move results to experiment

Next, several AI models are extracted from various real life authors from Kaggle. They are finetuned to perform well on the original dataset. For one, a random forest model is finetuned by searching hyperparameters settings\cite{Ipbyrne2023}. For another, a logistic regression model is finetuned by performing principal component analysis\cite{Prashant1112023}.

>>>>

After synthesizing the dataset, we used the fairness checker from \cite{yuan2024ensuring} to compute the fairness measures of any incoming AI model.

The fairness checker is an open-sourced public domain Python package that computes various fairness measures, such as those mentioned in Table~\ref{measures}.

The checker is designed to be user-friendly and agnostic to the underlying AI model. It is also designed to be easily extensible to accommodate new fairness measures.

The checker simply iterates through the given database $\db$ and computes the results based on some given predicates on the rows $r_i$s, and finally outputs the fairness measure values.

Protected groups $S$, predicted outcomes $\hat{Y}$, and ground truths $Y$ are all formulated as these predicates. These are straightforward logical boolean expressions. Specifically, they are given as Python functions that output boolean values.

For example, if the sensitive attribute is sex and the protected group is female, the protected group predicate would be $S := r_i(\text{``sex''}) = \text{``Female''}$. This can be easily implemented in Python as a comparison function.

The interpretation of the resulting fairness measure values is dependent on the third-party auditors. The auditors may have different thresholds $\epsilon$ for different fairness measures or different AI models.

To test the viability of our method, we compare the metrics computed from the synthetic dataset against those of the original dataset. We used various datasets with fairness concerns mentioned in \cite{pessach2022review}.

@@

Next, several AI models are extracted from various real-life authors from Kaggle. They are finetuned to perform well on the original dataset. For one, a random forest model is finetuned by searching hyperparameters settings\cite{Ipbyrne2023}. For another, a logistic regression model is finetuned by performing principal component analysis\cite{Prashant1112023}.
